---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = F
)
```

# modelimpact

<!-- badges: start -->
`r badger::badge_github_actions("PeerChristensen/modelimpact")`

<!-- badges: end -->

This package is intended to help data scientists and decision-makers understand the potential value of churn prediction models depending on how many customers are being targeted by a campaign.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` {r install}
# install.packages("devtools")
devtools::install_github("PeerChristensen/modelimpact")
```
## Functions and parameters

The first three functions aim to provide information about the business impact of using a model and targeting x % of the customer base. These functions accept the following arguments (required ones in
**bold**): <br>

-   **`x`** - a data frame <br>
-   `fixed_cost` - fixed costs (defaults to 0) <br>
-   `var_cost` - variable costs (defaults to 0) <br>
-   `tp_val` - true positive value (defaults to 0) <br>
-   **`prob_col`** - the variable containing target class probabilities
-   **`truth_col`** the variable containing the actual class

`profit_thresholds()` accepts the following arguments:

-   **`x`** - a data frame <br>
-   `var_cost` - variable costs <br>
-   `prob_accept` - Probability of offer being accepted. Defaults to 1. <br>
-   `tp_val` - The average value of a True Positive. `var_cost` is automatically subtracted. <br>
-   `fp_val` - The average cost of a False Positive. `var_cost` is automatically subtracted. <br>
-   `tn_val` - The average cost of a True Negatives <br>
-   `fn_val` - The average cost of a False Negatives <br>
-   **`prob_col`** - The column with probabilities of the event of interest <br>
-   **`truth_col`** - the column with the actual outcome/class. Possible values are 'Yes' and 'No'


```{r}
# Parameter settings
fixed_cost <- 1000
var_cost   <- 100
tp_val     <- 2000
```

## Costs and revenue

```{r data}
library(modelimpact)
library(tidyverse)
library(scales)

head(predictions)
```

```{r cost_rev}
cost_rev <- predictions %>%
  cost_revenue(
    fixed_cost = fixed_cost,
    var_cost   = var_cost,
    tp_val     = tp_val,
    prob_col   = Yes,
    truth_col  = Churn)

head(cost_rev)
```

```{r cost_rev_plot}
# functions for formatting plotting axes
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }

pcts <- function (x) { percent_format(scale=1)((x / max(x)) * 100) }
theme_set(theme_minimal())

cost_rev %>%
  ggplot() +
  geom_line(aes(row,cost_sum), colour ="black",linetype="dashed") + 
  geom_line(aes(row,cum_rev), colour = "darkred",size=1) + 
  scale_y_continuous(labels = ks) +
  scale_x_continuous(labels = pcts) +
  labs(x = "% targeted",y = "Costs & revenue")
```

## Profit

```{r profit}
profit_df <- predictions %>%
  profit(
    fixed_cost = fixed_cost,
    var_cost   = var_cost,
    tp_val     = tp_val,
    prob_col   = Yes,
    truth_col  = Churn)

head(profit_df)
```

```{r max-profit}
# max profit
max_profit <- profit_df %>% filter(profit == max(profit)) %>% select(row,pct,profit)

max_profit
```

```{r profit-plot}
profit_df %>%
  ggplot(aes(x=row,y=profit)) +
  geom_line(colour = "darkred",size=1) +
  scale_y_continuous(labels = ks) +
  geom_segment(x = max_profit$row, y= 0,xend=max_profit$row,
             yend = max_profit$profit, colour="black",linetype="dashed") +
  geom_hline(yintercept = 0,colour="black", linetype="dashed") +
  scale_x_continuous(labels = pcts) +
  labs(x = "% targeted",y = "Profit")
```

## Return on investment

```{r roi}
roi_df <- predictions %>%
  roi(
    fixed_cost = fixed_cost,
    var_cost   = var_cost,
    tp_val     = tp_val,
    prob_col   = Yes,
    truth_col  = Churn)

head(roi_df)
```

```{r roi-plot}
roi_df %>%
  ggplot(aes(x=row,y=roi)) +
  geom_hline(yintercept = 0,colour="black", linetype="dashed") +
  geom_line(colour = "darkred",size=1) +
  scale_x_continuous(labels = pcts) +
  labs(x = "% targeted",y = "ROI")
```

## Optimal threshold

```{r}
thresholds <- predictions %>%
  profit_thresholds(var_cost    = 100,
                    prob_accept = .7,
                    tp_val      = 2000,
                    fp_val      = 0,
                    tn_val      = 0,
                    fn_val      = -2000,
                    prob_col = Yes,
                    truth_col = Churn)

head(thresholds)
```
```{r}
optimal_threshold <- thresholds %>% filter(payoff == max(payoff))
optimal_threshold
```

```{r}
thresholds %>%
  ggplot(aes(x=threshold,y=payoff)) +
  geom_line(color="darkred",size = 1) +
  geom_hline(yintercept=0,linetype="dashed") +
  scale_y_continuous(labels = ks)
```

