---
title: "Beyond churn: applying modelimpact to other use cases"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Beyond churn: applying modelimpact to other use cases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  out.width = "100%"
)
has_ggplot2 <- requireNamespace("ggplot2", quietly = TRUE)
```

## The same engine, many problems

`modelimpact` was first written to reason about the business value of a
customer **churn** model, but nothing in the maths is churn-specific. Every
function does the same thing: it takes a column of predicted probabilities for
an event of interest and a column with the actual outcome, then translates the
model's behaviour into money using a handful of cost and value assumptions.

That makes the package directly usable for any **binary classification**
problem where acting on a prediction has a cost and being right (or wrong) has a
value, for example:

- **Fraud detection** — value of blocking a fraudulent transaction vs. the cost
  of reviewing a flagged one and the loss from fraud that slips through.
- **Credit default** — expected loss avoided vs. the cost of intervention.
- **Lead scoring and marketing response** — value of a converted lead vs. the
  cost of contacting it.
- **Upsell and cross-sell** — margin on an accepted offer vs. the cost of
  making it (use `prob_accept` to model take-up).
- **Predictive maintenance** — cost of an averted failure vs. the cost of an
  inspection.

The key is the `positive` argument, which every function accepts: it names the
value in your outcome column that identifies the event of interest. It defaults
to `"Yes"`, but it can be anything (`"fraud"`, `"default"`, `"converted"`, ...).

## A worked example: fraud detection

Suppose we have scored a batch of transactions with a fraud model. Each row is
a transaction with the model's predicted probability of fraud (`score`) and
whether it later turned out to be fraudulent (`fraud`). Fraud is rare, so the
classes are highly imbalanced.

```{r data}
library(modelimpact)
library(magrittr)  # for the %>% pipe

set.seed(42)
n <- 5000

# ~3% of transactions are actually fraudulent
is_fraud <- rbinom(n, size = 1, prob = 0.03)

# the model scores fraud higher than legitimate transactions, with overlap
score <- plogis(rnorm(n, mean = ifelse(is_fraud == 1, 1.5, -2.5), sd = 1.2))

transactions <- data.frame(
  score = score,
  fraud = ifelse(is_fraud == 1, "fraud", "legit")
)

head(transactions)
```

### The economics

We express the decision in monetary terms:

```{r econ}
review_cost  <- 8      # cost to manually review a flagged transaction
fraud_value  <- 400    # average value recovered when fraud is caught
missed_fraud <- -400   # average loss when fraud is not caught
```

### Cost and revenue as we review more transactions

Ranking transactions from most to least suspicious, how do the cumulative cost
of reviewing and the cumulative value recovered grow?

```{r cost-revenue, eval = has_ggplot2}
library(ggplot2)

transactions %>%
  cost_revenue(
    var_cost  = review_cost,
    tp_val    = fraud_value,
    prob_col  = score,
    truth_col = fraud,
    positive  = "fraud"
  ) %>%
  autoplot()
```

### How far down the ranked list should we review?

`profit()` shows the net value of reviewing the top *X %* of most suspicious
transactions, and `break_even()` / `impact_summary()` read off the key
operating points.

```{r profit, eval = has_ggplot2}
transactions %>%
  profit(
    var_cost  = review_cost,
    tp_val    = fraud_value,
    prob_col  = score,
    truth_col = fraud,
    positive  = "fraud"
  ) %>%
  autoplot()
```

```{r summary}
impact_summary(
  transactions,
  var_cost  = review_cost,
  tp_val    = fraud_value,
  prob_col  = score,
  truth_col = fraud,
  positive  = "fraud"
)
```

### Choosing a decision threshold

If the fraud model must instead produce an automatic block/allow decision,
`profit_thresholds()` sweeps every probability cutoff and scores the full
confusion matrix using a value for each cell. Here a missed fraud
(`fn_val`) is expensive, a false alarm (`fp_val`) merely costs a review, and
correct decisions on legitimate traffic are neutral.

```{r thresholds, eval = has_ggplot2}
transactions %>%
  profit_thresholds(
    var_cost  = review_cost,
    tp_val    = fraud_value,
    fp_val    = 0,
    tn_val    = 0,
    fn_val    = missed_fraud,
    prob_col  = score,
    truth_col = fraud,
    positive  = "fraud"
  ) %>%
  autoplot()
```

## Adapting to your own use case

To reuse any of these functions for a different problem, you only need to:

1. Point `prob_col` at your model's predicted-probability column and
   `truth_col` at the actual-outcome column.
2. Set `positive` to the label of the event you care about.
3. Translate your domain into the cost and value arguments (`var_cost`,
   `fixed_cost`, `tp_val`, and — for `profit_thresholds()` — `fp_val`,
   `tn_val`, `fn_val`), optionally using `prob_accept` when an action only
   incurs its cost if accepted.

The interpretation of every plot and summary then carries over unchanged.
